#Reproducible Research: Peer Assessment 1

##reading in the dataset and/or processing the data

```{r}
require(knitr)
knit2html("PA1_template.Rmd")

filePath<- getwd()
if(!file.exists("./PA1")){dir.create("./PA1")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip" 
download.file(fileUrl, destfile = "./Activity_monitoring_data.zip")

unzip(zipfile ="./Activity_monitoring_data.zip", exdir="./PA1")

variable.class<-c('numeric','character','numeric')
activity <- read.csv('./PA1/activity.csv',header=TRUE,na.strings='NA', colClasses = variable.class)
activity$date <- as.Date(activity$date, "%Y-%m-%d")

```

##Histogram of the total number of steps taken each day

```{r}

require(ggplot2)

totalSteps <- tapply(activity$steps, activity$date, FUN=sum, na.rm=TRUE)

qplot(totalSteps, binwidth= 1000, xlab="total number of steps taken each day", ylab="Frequecy", main="Fig1")

```

##Mean and median number of steps taken each day

```{r}
mean(totalSteps, na.rm=TRUE)
median(totalSteps, na.rm=TRUE)

```

##Time series plot of the average number of steps taken

```{r}

averages <- aggregate(steps ~ interval, data= activity, FUN = mean, na.rm=TRUE)

qplot(interval,steps , data= averages ,  geom= "line", xlab="5-Minute Interval", ylab="Average Number of Steps Taken", main="Fig2")

```

##The 5-minute interval that, on average, contains the maximum number of steps

```{r}


averages[which.max(averages$steps),1]


```

##Code to describe and show a strategy for imputing missing data

```{r}
missing<- is.na(activity$steps)
table(missing)

complete<-activity
complete$steps<- ifelse(is.na(complete$steps),averages$steps[match(complete$interval,averages$interval)],complete$steps)


```

##Histogram of the total number of steps taken each day after missing values are imputed

```{r}

fullstepsperday<- aggregate(complete$steps, list(complete$date), FUN=sum)
qplot(fullstepsperday$x, binwidth= 5000, xlab="Steps Per Day", ylab="Frequecy", main="Fig3")

```

##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}

weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')

activity$wDay <- factor((weekdays(activity$date) %in% weekdays),levels=c(FALSE, TRUE), labels=c('weekend','weekday'))

averages2 <- aggregate(steps ~ interval + wDay , data= activity, FUN = mean, na.rm=TRUE) 

ggplot(averages2, aes(x=interval, y=steps, color = wDay)) + geom_line() + facet_wrap(~wDay, ncol = 1, nrow=2)+ ggtitle("Fig4")


```